abxD01 Paper

### 2015-04-04 PAPER.Rmd
Add rf code to text

```{r random_forest}

}

***The composition of the microbiota on day 0 is predictive of *C. difficile* colonization levels on day 1.*** The three sets of experiments demonstrated that in certain contexts individual OTUs could be associated with *C. difficile* colonization, but in other contexts they were not. For example, the OTU affiliated with *Escherichia* (OTU 4) was associated with high levels of colonization in the ampicillin, metronidazole, and cefoperazone-treated mice; however, it was not detected in the streptomycin-treated mice although *C. difficile was able to colonize those mice to high levels. This suggests that colonization is a phenotype that is driven by multiple populations that act independently and possibly in concert to resist colonization. Therefore, we used a regression-based random forest machine learning algorithm to predict the level of *C. difficile* colonization observed in the three sets of experiments based on the composition of the microbiota at the time of challenge. The model explained XXX98.0XXX% of the variation in the observed *C. difficile* colonization levels (Figure 5).

Many of the OTUs that contributed the most to the quality of the fit included members of the *Porphyromonadaceae*, *Alistipes*, *Lachnospiraceae*, *Lactobacillus*, and *Escherichia*. These results further validate the observations from the correlation-based analysis. Together these results suggest that colonization resistance is likely conferred by *porphyromonadaceae*, *Alistipes*, and *lachnospiraceae* and a loss in these bacteria, concurrently with a gain of *Escherichia*, can result in increased susceptibility to infection. <needs help>Interestingly, the relationship between these OTUs and colonization resistance was not a simple dichotomous or linear relationship. For example, among the streptomycin-treated mice which altogether lacked the *Enterobacteriaceae*, *Porphyromonadaceae* (OTU 3) and *Lachnospiraceae* (OTU 39), were the two most important features in accurately predicting subsequent *C. difficile* levels. In fact, a model built with only these two features was able to account for 79% of the variance observed. </needs help>

>>>>

```{r random_forest}
data <- read.table(file="data/process/random_forest.data")
n_full_features <- data[1,1]
rsq_full_features <- data[1,2]

n_part_features <- data[2,1]
rsq_part_features <- data[2,2]
```

***The composition of the microbiota on day 0 is predictive of *C. difficile* colonization levels on day 1.*** The three sets of experiments demonstrated that in certain contexts individual OTUs could be associated with *C. difficile* colonization, but in other contexts they were not. For example, the OTU affiliated with *Escherichia* (OTU 4) was associated with high levels of colonization in the ampicillin, metronidazole, and cefoperazone-treated mice; however, it was not detected in the streptomycin-treated mice although *C. difficile was able to colonize those mice to high levels. This suggests that colonization is a phenotype that is driven by multiple populations that act independently and possibly in concert to resist colonization. Therefore, we used a regression-based random forest machine learning algorithm to predict the level of *C. difficile* colonization observed in the three sets of experiments based on the composition of the microbiota at the time of challenge. The model explained `r round(rsq_full_features*100, 1)`% of the variation in the observed *C. difficile* colonization levels (Figure 5). Noticing a natural jump following the twelfth OTU in the percent increase in the mean squared error when each OTU was removed, we regenerated the model using only the top 12 OTUs. This model explained `r round(rsq_part_features*100, 1)`% of the variation in the observed *C. difficile* colonization levels.

Many of the OTUs that contributed the most to the quality of the fit included members of the *Porphyromonadaceae*, *Alistipes*, *Lachnospiraceae*, *Lactobacillus*, and *Escherichia*. These results further validate the observations from the correlation-based analysis. Together these results suggest that colonization resistance is likely conferred by *porphyromonadaceae*, *Alistipes*, and *lachnospiraceae* and a loss in these bacteria, concurrently with a gain of *Escherichia*, can result in increased susceptibility to infection. <needs help>Interestingly, the relationship between these OTUs and colonization resistance was not a simple dichotomous or linear relationship. For example, among the streptomycin-treated mice which altogether lacked the *Enterobacteriaceae*, *Porphyromonadaceae* (OTU 3) and *Lachnospiraceae* (OTU 39), were the two most important features in accurately predicting subsequent *C. difficile* levels</needs help>
